---
title: "Data Manipulation with dplyr"
output: github_document
date: "2022-09-22"
---

%>% is called a "pipe", read as "and then"

Once you've imported data, you're going to need to so some cleaning up.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  fig.width = 6
)
```


```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you'd like to remove: telling it you want everything except the pups_survive. Can drop more variables by doing , -variable name. 

```{r}
select(litters_data, -pups_survive)
```

In this process of selecting columns you are gonna use, you want to rename your variables. So with select you can select the columns and rename them in one step:

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

If all you want to do is rename something, you can use `rename` instead of `select`. This will rename the variables you care about, and keep everything else. Here have full data set again just with these 2 new names. With `select` you don't keep all the other variables in the dataframe (they get dropped if not renamed); don't drop other variables with `rename`. Can't do anything with this dataset, just trying to show us what this function can do.

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
 
When using `select` you can do some dif things (like above) but there are smarter ways. Helper functions with `select`: ?select_helpers. I use starts_with(), ends_with(), and contains() often, especially when there variables are named with suffixes or other standard patterns:

Can use `starts_with()` function to grab the data named similarly...

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
```

Sometimes just want to change variable you see first. This function reorganizes the columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

`select` lets us select columns, but maybe we want to select rows. Rows representing one observation, or in this case a litter of pups. Use `filter`. You will often filter using comparison operators (`>`, `>=', `<`, `<=`, `==`, and `!=`). `==` Is when you want something equal to x value. Need two == because it is not an assignment (X is now Y) but when is X the exact same value as some number or whatever. You may also use %in% to detect if values appear in a set, and `is.na()` to find missing values. The results of comparisons are logical – the statement is `TRUE` or `FALSE` depending on the values you compare – and can be combined with other comparisons using the logical operators `&` and `|`, or negated using `!`.

If true will get filtered into new dataset, if false won't go. Sometimes need to combine multiple conditions with "ands" & and "ors" | 

* `gd_of_birth == 20`
* `pups_born_alive >= 2`
* `pups_survive != 4`
* `!(pups_survive == 4) & (gd_of_birth ==20))`
* `group %in% c("Con7", "Con8")`
* `group == "Con7" & gd_of_birth == 20`

```{r}
filter(litters_data, gd_of_birth == 20)
```

```{r}
filter(litters_data, group == "Con7" & gd_of_birth == 20)
```

A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using drop_na from the tidyr package:

drop_na(litters_data) will remove any row with a missing value
drop_na(litters_data, wt_increase) will remove rows for which wt_increase is missing.


