---
title: "Data Manipulation with dplyr"
output: github_document
date: "2022-09-22"
---

%>% is called a "pipe", read as "and then"

Once you've imported data, you're going to need to so some cleaning up.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  fig.width = 6
)
```


```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you'd like to remove: telling it you want everything except the pups_survive. Can drop more variables by doing , -variable name. 

```{r}
select(litters_data, -pups_survive)
```

In this process of selecting columns you are gonna use, you want to rename your variables. So with select you can select the columns and rename them in one step:

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```





