---
title: "Data Import"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Data Import: CSVs

Let's import data using the `readr` package

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
```{r}
litters_df

head(litters_df)
tail(litters_df)
```

```{r, eval = FALSE}
view(litters_df)
```

Skimr package :: means use the skim function
```{r}
skimr::skim(litters_df)
```

`read_csv` options ..

How to say the data here is missing and how to skip rows with `skip =`
```{r, eval = FALSE}
read_csv("data/FAS_litters.csv", na = c("", "NA"))
```

## OTHER FILE FORMATS

We need to read in an excel spreadsheet...

```{r}
mlb_df = read_excel("data/mlb11.xlsx")
```

```{r, eval = FALSE}
View(mlb_df)
```

read_excel("data/mlb11.xlsx")
```{r}
lotr_words = read_excel(
  "data/LotR_Words.xlsx",
  range = "B3:D6")
```

```{r,eval = FALSE}
View(lotr_words)
```

## Still more formats...

Read in a SAS dataset

```{r, eval = FALSE}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
View(pulse_df)
```

## Data Export

Data you want to write and where you want to put it 
```{r}
write_csv(lotr_words, file = "data/lotr_words_df.csv")
```

Don't use read.csv only use read_csv

Can tell R to ignore things in the .gitignore tab by just typing in name of thing you want to ignore. `data/lotr_words_df.csv` for example

If you want to discard changes you made, in the commit window you can delete the change but you can't take it back so be fr.

Reminder:
tidyverse works with CSVs
readxl works with excel sheets
haven works with SAS, or other file types 
