---
title: "Tidying_data_rmd"
output: github_document
date: "2022-09-27"
---

# Slide Notes

*Columns are variables, rows are observations, cells are values
Not all data can be tidy though...

*Nontidy example is when columns aren't variables, when they are values like 
"treatment a" "treatment b" instead of individual columns like "treatment" is the variable in the columns, result is another column and the treatments a or b become the values in the cells.

*Tidying data can bring 3 separate tables together into 1

*Relational data: 2 excel spreadsheets with 2 data need to combine together. Subject 1 and outcome in one table and second table has something else. If table has NAs can drop those observations or ...

## Combining datasets
*`inner_join` only takes observations in both of the tables, so things not in both get dropped
*`left_join` drops anything in the right that not contained in left. Puts in NA where there is no info for right.
*`right_join` drops anything in the left that doesn't match/contained in right
*`full_join` takes every observation from both but puts NAs where there is not a value contained in both. Nothing dropped, just NA listed. 

##For tidying single tables 
`pivot_longer` changes data from wide format (cuz spread out wide) to long format bc long like less columns more observations 
`separate` separates columns that have a ton of observations in one column 
For untidying to `pivot wider` to do opposite of `pivot longer`
To combine multiple tables 
`*_join` = * = left, right, or full join 
`anti_join` shows you whats been dropped 

```{r}
library(tidyverse)

options(tibble.print_min = 5)
```

```{r}
pulse_df = haven::read_sas('data/public_pulse_data.sas7bdat')

View(pulse_df)

```

```{r}

pulse_df = haven::read_sas('data/public_pulse_data.sas7bdat') %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = 'visit',
    values_to = 'bdi',
    names_prefix = 'bdi_score_'
  ) %>%
  mutate(
    visit = replace(visit, visit == 'bl', '00n'),
    visit = factor(visit)
  ) %>%
  arrange(id, visit)

```

# Learning Assessment
```{r}

litters_df = read_csv('data/FAS_litters.csv') %>%
  janitor::clean_names() %>%
  select(litter_number, ends_with('weight')) %>%
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = 'gd',
    values_to = 'weight'
    ) %>%
  mutate(gd = recode(gd, 'gd0_weight' = 0, 'gd18_weight' = 18))
```

# Turn Long Data Wide

```{r}
analysis_results = tibble(
  group = c('treatment', 'treatment', 'placebo', 'placebo'),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_results_wide = pivot_wider(
  analysis_results,
  names_from = 'time',
  values_from = 'mean'
)
```

# Binding Rows

```{r}
fellowship_ring = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "B3:D6") %>%
  mutate(movie = 'fellowship_ring')

two_towers = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "F3:H6") %>%
  mutate(movie = 'two_towers')

return_king = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "J3:L6") %>%
  mutate(movie = 'return_king')

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = 'gender',
    values_to = 'words'
  ) %>%
  mutate(race = str_to_lower(race))
```

# Joining Data Frames

```{r}
pups =
  read_csv('data/FAS_pups.csv') %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = 'male', `2` = 'female'))

litter = 
  read_csv('data/FAS_litters.csv') %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose)
    )

fas_data = 
  left_join(pups, litter, by = "litter_number")
```


